### 2017-09-25 analysis/Simulation_generic_Cormotif.Rmd
wflow_publish("Simulation_generic_Cormotif.Rmd")

---
title: "Simulation for generic Cormotif model"
author: "First Last"
date: "9/24/2017"
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

<!-- Add your analysis here -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,eval=TRUE)
```

## Simulation

We set unit number $n = 5000$, and study number $R = 4$. We set every observation has the same standard error, $s_{jr}=1$. That is $\hat\beta_{jr}|\beta_{jr}\sim N(\beta_{jr};0,1)$. The 5000 units come from 4 patterns ($K=4$): 2000 units have zero effects in all four studies, that is $\beta_{jr}=0$, for $r = 1,2,3,4$; 1000 units have effect $\beta_{jr}\sim N(0,10^2)$ for $r =1, 2$ and $\beta_{jr}=0$ for $r =3, 4$; 1000 units have effect $\beta_{jr}\sim N(0,10^2)$ for $r =2, 3$ and $\beta_{jr}=0$ for $r =1, 4$; 1000 unitss have effect $\beta_{jr}\sim N(0,10^2)$ for $r =1, 2,3,4$.

```{r,cache=TRUE}
source('~/Git/cormotif/code/function.R')
K = 4
D = 4
G = 5000
sigma2 = c(100,100,100,100)
pi0 = c(0.2,0.2,0.2,0.4)
q0 = matrix(0,nrow=K,ncol=D)
q0[1,] = c(1,1,1,1)
q0[2,] = c(1,1,0,0)
q0[3,] = c(0,1,1,0)
q0[4,] = c(0,0,0,0)
X = matrix(0,nrow=G,ncol=D)
rows = rep(1:K,times=pi0*G)
A = q0[rows,]
set.seed(111)
for(g in 1:G){
  for(d in 1:D){
    if(A[g,d]==1){
      X[g,d] = rnorm(1,0,sqrt(1+sigma2[d]))
    } else{
      X[g,d] = rnorm(1,0,1)
    }
  }
}
betahat=X
sebetahat=matrix(1,nrow=G,ncol=D)
```

```{r,cache=TRUE}
fit = generic.cormotif(betahat,sebetahat,K=1:5)
```

After fitting the generic cormotif model, we can check the BIC values obtained by all cluster numbers:

```{r}
fit$bic
```

$K = 4$ has the smallest BIC. The standard deviation $\sigma_0,\sigma_1,\ldots,\sigma_L$ built by our method is

```{r}
fit$mixsd
```

And the estimated $\pi$ for $K = 4$ is

```{r}
fit$bestmotif$pi
```

Then we could check the estimation for $g$, i.e. the values of $w$. The estimated $w$ for studies 1 is

```{r}
fit$bestmotif$W[[1]]
```

The estimated $w$ for studies 2 is

```{r}
fit$bestmotif$W[[2]]
```

The estimated $w$ for studies 3 is

```{r}
fit$bestmotif$W[[3]]
```

The estimated $w$ for studies 4 is

```{r}
fit$bestmotif$W[[4]]
```

All of them matches well with our setting, which verifies our algorithm.

Also, we could compute the lfdr:

```{r}
head(cbind(fit$result[[1]]$lfdr,fit$result[[2]]$lfdr,fit$result[[3]]$lfdr,fit$result[[4]]$lfdr))
```

And lfsr:

```{r}
head(cbind(fit$result[[1]]$lfsr,fit$result[[2]]$lfsr,fit$result[[3]]$lfsr,fit$result[[4]]$lfsr))
```

## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```