### 2017-09-29 analysis/Trend_of_loglike.Rmd
wflow_publish("Trend_of_loglike.Rmd")

---
title: "Trend of Loglike"
author: "Zhiwei Ma"
date: "9/24/2017"
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

<!-- Add your analysis here -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,eval=FALSE)
```

## Simulation 1

We set unit number $n = 5000$, and study number $R = 4$. We set every observation has the same standard error, $s_{jr}=1$. That is $\hat\beta_{jr}|\beta_{jr}\sim N(\beta_{jr};0,1)$. The 5000 units come from 4 patterns ($K=4$): 2500 units have zero effects in all four studies, that is $\beta_{jr}=0$, for $r = 1,2,3,4$; 1000 units have effect $\beta_{jr}\sim N(0,4^2)$ for $r =1, 2$ and $\beta_{jr}=0$ for $r =3, 4$; 1000 units have effect $\beta_{jr}\sim N(0,4^2)$ for $r =3, 4$ and $\beta_{jr}=0$ for $r = 1, 2$; 500 unitss have effect $\beta_{jr}\sim N(0,4^2)$ for $r =1, 2,3,4$.

To fit the model, set $K = 1:10$.

```{r,cache=TRUE}
source('../code/function.R')
# Simulate data
K = 4
D = 4
G = 5000
sigma2 = c(16,16,16,16)
pi0 = c(0.1,0.2,0.2,0.5)
q0 = matrix(0,nrow=K,ncol=D)
q0[1,] = c(1,1,1,1)
q0[2,] = c(1,1,0,0)
q0[3,] = c(0,0,1,1)
q0[4,] = c(0,0,0,0)
X = matrix(0,nrow=G,ncol=D)
rows = rep(1:K,times=pi0*G)
A = q0[rows,]
set.seed(222)
for(g in 1:G){
  for(d in 1:D){
    if(A[g,d]==1){
      X[g,d] = rnorm(1,0,sqrt(1+sigma2[d]))
    } else{
      X[g,d] = rnorm(1,0,1)
    }
  }
}
betahat=X
sebetahat=matrix(1,nrow=G,ncol=D)
```

```{r,cache=TRUE}
#fit the model
fit1 = generic.cormotif(betahat,sebetahat,K=1:10,mess=FALSE)
```

The estimated $\pi$ for $K=4$

```{r}
fit1$allmotif[[4]]$pi
```

we can check the BIC and AIC values obtained by all cluster numbers:

```{r}
fit1$bic
```

The AIC values:

```{r}
fit1$aic
```

The loglikelihood values:

```{r}
fit1$loglike
```

And plot

```{r}
plot(fit1$loglike[,2],type = "l",xlab = "K",ylab = "loglike")
```

## Simulation 2

With the same setting as simulation 1, just change $\beta_{jr}\sim N(0,10^2)$, we check the same process.

To fit the model, set $K = 1:10$.

```{r,cache=TRUE}
#source('../code/function.R')
# Simulate data
K = 4
D = 4
G = 5000
sigma2 = c(100,100,100,100)
pi0 = c(0.1,0.2,0.2,0.5)
q0 = matrix(0,nrow=K,ncol=D)
q0[1,] = c(1,1,1,1)
q0[2,] = c(1,1,0,0)
q0[3,] = c(0,0,1,1)
q0[4,] = c(0,0,0,0)
X = matrix(0,nrow=G,ncol=D)
rows = rep(1:K,times=pi0*G)
A = q0[rows,]
set.seed(22)
for(g in 1:G){
  for(d in 1:D){
    if(A[g,d]==1){
      X[g,d] = rnorm(1,0,sqrt(1+sigma2[d]))
    } else{
      X[g,d] = rnorm(1,0,1)
    }
  }
}
betahat=X
sebetahat=matrix(1,nrow=G,ncol=D)
```

```{r,cache=TRUE}
#fit the model
fit2 = generic.cormotif(betahat,sebetahat,K=1:10,mess=FALSE)
```

The estimated $\pi$ for $K=4$

```{r}
fit2$allmotif[[4]]$pi
```

we can check the BIC and AIC values obtained by all cluster numbers:

```{r}
fit2$bic
```

The AIC values:

```{r}
fit2$aic
```

The loglikelihood values:

```{r}
fit2$loglike
```

And plot

```{r}
plot(fit2$loglike[,2],type = "l",xlab = "K",ylab = "loglike")
```

## Simulation 3

With the same setting as simulation 1, just change the proportion of 4 patterns to 1250:1250:1250:1250, we check the same process.

To fit the model, set $K = 1:10$.

```{r,cache=TRUE}
#source('../code/function.R')
# Simulate data
K = 4
D = 4
G = 5000
sigma2 = c(16,16,16,16)
pi0 = c(0.25,0.25,0.25,0.25)
q0 = matrix(0,nrow=K,ncol=D)
q0[1,] = c(1,1,1,1)
q0[2,] = c(1,1,0,0)
q0[3,] = c(0,0,1,1)
q0[4,] = c(0,0,0,0)
X = matrix(0,nrow=G,ncol=D)
rows = rep(1:K,times=pi0*G)
A = q0[rows,]
set.seed(333)
for(g in 1:G){
  for(d in 1:D){
    if(A[g,d]==1){
      X[g,d] = rnorm(1,0,sqrt(1+sigma2[d]))
    } else{
      X[g,d] = rnorm(1,0,1)
    }
  }
}
betahat=X
sebetahat=matrix(1,nrow=G,ncol=D)
```

```{r,cache=TRUE}
#fit the model
fit3 = generic.cormotif(betahat,sebetahat,K=1:10,mess=FALSE)
```

The estimated $\pi$ for $K=4$

```{r}
fit3$allmotif[[4]]$pi
```

we can check the BIC and AIC values obtained by all cluster numbers:

```{r}
fit3$bic
```

The AIC values:

```{r}
fit3$aic
```

The loglikelihood values:

```{r}
fit3$loglike
```

And plot

```{r}
plot(fit3$loglike[,2],type = "l",xlab = "K",ylab = "loglike")
```

## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```