### 2018-01-12 analysis/MTB_Expression_Analysis.Rmd
wflow_publish("MTB_Expression_Analysis.Rmd")

---
title: "MTB Expression Analysis"
author: "Zhiwei Ma"
date: 2018-01-12
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,eval=FALSE)
```

<!-- Add your analysis here -->

## Construction of Z-scores

For each bacterial infection at certain timepoint, we have six controls and six treatments, denoted as $X_1,X_2,X_3,X_4,X_5,X_6$ and $Y_1,Y_2,Y_3,Y_4,Y_5,Y_6$. Then the estimated effect size and variance are defined below:
$$
\hat\beta = \bar Y-\bar X; \\
\hat s = \sqrt{\hat\sigma_x^2/6+\hat\sigma_y^2/6.}
$$
Here $\bar X$, $\bar Y$ are sample means, and $\hat\sigma_x$, $\hat\sigma_y$ are sample 
variances.

## Motifs

At 18 hours post-infection, There are eight bacterial infections: "Rv.18", "Rv+.18", "GC.18", "BCG.18", "Smeg.18", "Yers.18", "Salm.18", "Staph.18". Here we set $K=5$, the same like [Blischak, et al, 2015](https://www.nature.com/articles/srep16882). The motifs are shown below.

```{r}
source('../code/function.R')
betahat = read.table("../data/betahat.txt", header = TRUE, sep = "\t",
               stringsAsFactors = FALSE)
sebetahat = read.table("../data/sebetahat.txt", header = TRUE, sep = "\t",
               stringsAsFactors = FALSE)
z18_k5 = generic.cormotif(betahat[,9:16],sebetahat[,9:16],K=5,mess=FALSE)
plotMotif(z18_k5$bestmotif)
```

## Detection of eQTLs

Download the 198 response eQTLs from [Barreiro et al, 2012](https://www.ncbi.nlm.nih.gov/pubmed/22233810) (179 out of the 198 were also expressed in our study). Apply our method, we can compute the $lfsr$ (local false sign rate) for each effect. Set the thresholding at 0.01, we detect 59 differential expression out of 179 genes.

```{r}
dat_cpm <- read.table("../data/table-s1.txt", header = TRUE, sep = "\t",
                      stringsAsFactors = FALSE)
# The first two columns are gene information
rownames(dat_cpm) <- dat_cpm$id
gene_names <- dat_cpm$name
names(gene_names) <- dat_cpm$id

id = rownames(dat_cpm)
lfsr18 = z18_k5$lfsr[[1]][,1] 
list = id[lfsr18<0.01]
a = intersect(S3[,2],list)
print(gene_names[a])
```

## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```