### 2018-01-29 analysis/Simulation_paper.Rmd
a

---
title: "Simulation in Paper"
author: "Zhiwei Ma"
date: 2018-01-28
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,eval=TRUE)
```

<!-- Add your analysis here -->

## Results

### Simulation 1

We set unit number $n = 10000$, and study number $R = 4$. We set every observation to be the same standard error, $s_{jr}=1$. That is $\hat\beta_{jr}|\beta_{jr}\sim N(\beta_{jr};0,1)$. The 10000 units come from 4 patterns ($K=4$): 9000 units have zero effects in all four studies, that is $\beta_{jr}=0$, for $r = 1,2,3,4$; 400 units have effect $\beta_{jr}\sim N(0,4^2)$ for $r =1, 2$ and $\beta_{jr}=0$ for $r =3, 4$; 400 units have effect $\beta_{jr}\sim N(0,4^2)$ for $r =2, 3$ and $\beta_{jr}=0$ for $r = 1, 4$; 200 unitss have effect $\beta_{jr}\sim N(0,4^2)$ for $r =1,2,3,4$.

To fit the model, set $K = 1:10$.

```{r}
source('../code/function.R')
K = 4
D = 4
G = 10000
sigma2 = c(16,16,16,16)
pi0 = c(0.02,0.04,0.04,0.9)
q0 = matrix(0,nrow=K,ncol=D)
q0[1,] = c(1,1,1,1)
q0[2,] = c(1,1,0,0)
q0[3,] = c(0,1,1,0)
q0[4,] = c(0,0,0,0)
X = matrix(0,nrow=G,ncol=D)
Y = matrix(0,nrow=G,ncol=D)
rows = rep(1:K,times=pi0*G)
A = q0[rows,]
set.seed(111)
for(g in 1:G){
  for(d in 1:D){
    if(A[g,d]==1){
      X[g,d] = rnorm(1,0,sqrt(sigma2[d]))
    } else{
      X[g,d] = 0
    }
  }
}
beta = X
sebetahat=matrix(1,nrow=G,ncol=D)
betahat=matrix(10,nrow=G,ncol=D)
for(g in 1:G){
  for(d in 1:D){
    betahat[g,d] = rnorm(1,beta[g,d],sebetahat[g,d])
  }
}

# fit the model
fit1 = generic.cormotif(betahat,sebetahat,K=1:10,mess=FALSE)
```

The loglikelihood value plot:

```{r}
plot(fit1$loglike[,2],type = "l",xlab = "K",ylab = "loglike")
```

The RRMSE for K = 1:10 is

```{r}
RRMSE = rep(0,10)
for(i in 1:10){
  RRMSE[i] = sqrt(mean((fit1$post_mean[[i]]-beta)^2)/mean((betahat-beta)^2))
}
plot(RRMSE,type="l",xlab="K",ylab="RRMSE")
```

Notice that when $K=1$, it is equvalent to applying ash to each study separately, and $K = 4$ is the true model.

The plot for lfsr

```{r}
plot(as.vector(fit1$lfsr[[4]][,1]),ylim=c(-0.1,1),type="l",xlab="Index",ylab="lfsr",col=1)
lines(as.vector(fit1$lfsr[[10]][,1]-fit1$lfsr[[4]][,1]),lty=2,col=2)
legend('right',legend=c("lfsr K=4","diff K=4,10"),lty =c(1,2),col=1:2)
```

Changing the threshold for lfsr, we plot the ROC curve for $K=1,2,4,10$.

```{r}
L = 10000
thres = seq(0.01,1,length=L)
fpr = rep(0,L)
tpr = rep(0,L)
for(i in 1:L){
  nT = sum(beta != 0)
  nCSS = sum((fit1$lfsr[[1]]<=thres[i])&(fit1$post_mean[[1]]*beta>0))
  nN = sum(beta == 0)
  nNS = sum((fit1$lfsr[[1]]<=thres[i])&(beta == 0))
  tpr[i] = nCSS/nT
  fpr[i] = nNS/nN
}
plot(fpr,tpr,type="l",xlab="False Positive Rate",ylab="True Positive Rate",col=1)
for(i in 1:L){
  nT = sum(beta != 0)
  nCSS = sum((fit1$lfsr[[2]]<=thres[i])&(fit1$post_mean[[2]]*beta>0))
  nN = sum(beta == 0)
  nNS = sum((fit1$lfsr[[2]]<=thres[i])&(beta == 0))
  tpr[i] = nCSS/nT
  fpr[i] = nNS/nN
}
lines(fpr,tpr,col=2)
for(i in 1:L){
  nT = sum(beta != 0)
  nCSS = sum((fit1$lfsr[[4]]<=thres[i])&(fit1$post_mean[[4]]*beta>0))
  nN = sum(beta == 0)
  nNS = sum((fit1$lfsr[[4]]<=thres[i])&(beta == 0))
  tpr[i] = nCSS/nT
  fpr[i] = nNS/nN
}
lines(fpr,tpr,col=3)
for(i in 1:L){
  nT = sum(beta != 0)
  nCSS = sum((fit1$lfsr[[10]]<=thres[i])&(fit1$post_mean[[10]]*beta>0))
  nN = sum(beta == 0)
  nNS = sum((fit1$lfsr[[10]]<=thres[i])&(beta == 0))
  tpr[i] = nCSS/nT
  fpr[i] = nNS/nN
}
lines(fpr,tpr,col=4)
legend('bottomright',legend=c("K=1","K=2","K=4","K=10"),lty =1,col=1:4)
```

## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```